---
title: "CatchUp_1R"
output: html_document
date: "`r format(Sys.time(), '%a %b %d %X %Y')`"
author: RZ

note: run catchup_1.cpp first to create data
---
```{r,include=FALSE}
rm(list=ls())
gc()
library(Rcpp)
options(warn=-1)
time.start<-proc.time()
```
Reference article is [Analysis of hash-rate double spending](https://arxiv.org/pdf/1402.2009.pdf)

We simulate the situation where we have:

a. one **attacker** and one (honest) **miner**
b. a number of **chains** (of **blocks**)
c. The **hashing power** for the miner is given by miner.hashpower in the range: 1,2,... 100. The hashing power of the attacker is: *attacker.hashpower = 100 - miner.hashpower*

The attacker has a **backlog** of a certain number of blocks and has to **catch up** with the miner (backlog=0). We say: the attacker 'wins'.
Since we simulate we have to define:

d. a point in time to **stop trying to catch up** (max.time)
e. a number of **iterations** (iterations) for the simulation. Increasing this number gives a better estimation of the attacker's expected win chance due to [Law of The Large Numbers](https://en.wikipedia.org/wiki/Law_of_large_numbers).

```{r,echo=FALSE}
#Earlier we found that
#$\sum_{X_1=0}^{[t/2-B]} \sum_{X_2=X_1+B+1}^{t-X1} t!/X_1! * X_2! * (t-X_1-X_2) * p_1^X_1 * p_2^X_2 * p_3^(t-X_1-X_2)$
```

```{r simulate, echo=FALSE,evaluate=FALSE}
simulate<-function(chains,backlog,max.time,miner.hashpower,iterations){
  attacker.hashpower<-100 - miner.hashpower
  miner.hashpower<-miner.hashpower/100
  attacker.hashpower<-attacker.hashpower/100
  catchup.times<-c()
  catchup.succeeded.count<-catchup.fail.count<-0

  for (ix.iteration in seq_len(iterations)){
    situation.backlogs<-rep(backlog,chains)
    miner.assign.chain<-sample(seq_len(chains),max.time,replace=TRUE) # chain for generated block miner
    attacker.assign.chain<-sample(seq_len(chains),max.time,replace=TRUE) # chain for generated block attacker
    miner.will.generate<-(runif(max.time)<=miner.hashpower) # will the miner generate?
    attacker.will.generate<-(runif(max.time)<=attacker.hashpower) #will the attacker generate?

    for (ix.time in seq.int(max.time)){
          situation.backlogs[miner.assign.chain[ix.time]]<-
            situation.backlogs[miner.assign.chain[ix.time]]+miner.will.generate[ix.time]
          situation.backlogs[attacker.assign.chain[ix.time]]<-
            situation.backlogs[attacker.assign.chain[ix.time]]-attacker.will.generate[ix.time]
        catchup.succeed<-any(situation.backlogs<=0)
        if (catchup.succeed){
          catchup.succeeded.count<-catchup.succeeded.count+1
          catchup.times<-c(catchup.times,ix.time)
          break}
    }
    catchup.fail.count<-catchup.fail.count+1
    if (catchup.fail.count>catchup.fail.limit) {break}
  }
  #return(list(catchup.succeeded.count,mean(catchup.times)))
  return(c(catchup.succeeded.count,mean(catchup.times)))
 
}
```

```{r,include=FALSE}
test.mode<-FALSE # set to FALSE for large parameters
if (test.mode) {
  BL<-c(2,3)
  HPM<-c(50,52)} else {
  BL<-seq(2,7)
  HPM<-seq(50,95,5)  
}
```


```{r set parameters of the procedure,echo=TRUE,warning=FALSE,fig.width=8,fig.height=6}
chains<-1
max.time<-1000
iterations<-30

# procedure settings
table.only<-TRUE # set this to FALSE to also show graphics
catchup.fail.limit<-iterations%/%2 # used for speeding up iteration: if number of failed catchup exceeds catchup.fail.limit, we assume/expect that we see no more catchups and therefore we stop iteration
```


```{r info backlogs and haspowers miner,echo=FALSE}
cat('backlogs in simulations:',BL)
cat('hashpower miner in simulations:',HPM)
```

```{r c++ procedure,include=FALSE,echo=FALSE}
chains <- 2
max.time <- 10000
iterations <- 1000
test.mode<-FALSE # set to FALSE for large parameters
if (test.mode) {
  BL<-c(2,3)
  HPM<-c(50,52)} else {
  BL<-seq(2,7)
  HPM<-seq(50,98,2)  
}
table.results<-matrix(nrow=length(HPM),ncol=length(BL),dimnames=list(sort(100-HPM),BL))
for (miner.hashpower in HPM){
  for (backlog in BL){
    #table.results[as.character(100-miner.hashpower),as.character(backlog)]<-Csimulate(chains,backlog,max.time,miner.hashpower,iterations)} #save result in matrix
  }
}
#print(table.results)

```

```{r procedure,echo=FALSE,eval=FALSE,warning=FALSE,fig.width=8,fig.height=6}
table.results<-matrix(nrow=length(HPM),ncol=length(BL),dimnames=list(sort(100-HPM),BL))
kleuren<-rep(c('red','green','blue','black','pink','purple','orange','grey'),3)

for (miner.hashpower in HPM){

  H.all<-list()
  ix.kleur<-0
  win.chances.all<-breaks.all<-c()
  
  for (backlog in BL){

    par(mfrow=c(2,2))
    par(new=FALSE)
    ix.kleur<-ix.kleur+1
    min.time <- backlog # note that the attacker can not win by definition if time<backlog
    rep.factor <- max.time-min.time+1 # technical parameter to drive the lengths of the vectors for mcmapply
    res<-mcmapply(simulate,rep(chains,rep.factor),rep(backlog,rep.factor),seq(min.time,max.time),rep(miner.hashpower,rep.factor),rep(iterations,rep.factor),SIMPLIFY = 'array')
    wins<-res[1,]
    catch.up<-res[2,]
    
    if (all(is.na(catch.up))) {
      win.chances.last<-0
    } else {
      # er is tenminste een keer een catchup geweest
          
      #PLOT 1

      H<-hist(catch.up,breaks=50,main='HISTOGRAM CATCHUP TIMES',xlab='time',ylab='freq',col='red',plot=!table.only)
      win.chances<-cumsum(H$counts)/(rep.factor)
      win.chances.last<-win.chances[length(win.chances)]

      if (!table.only){

        H.all<-c(H.all,list(miner.hashpower,backlog,H$breaks[2:length(H$breaks)],win.chances,kleuren[ix.kleur],sum(wins))) #save info
    
        #PLOT 2
        plot(H$breaks[2:length(H$breaks)],win.chances,main='CUM CHANCE CATCHUP TIMES',xlab='time',ylab='cum chance',col='red',type='h')
        abline(h = seq(0.1,1,.1), lty = 2)

        breaks.last<-H$breaks[length(H$breaks)]
        breaks.all<-c(breaks.all,breaks.last) #for overview usage
    
        #PLOT 3
        plot(H$breaks[2:length(H$breaks)],win.chances,main='WIN CHANCES',xlab='time',ylab='win.chances',col='green',type='s')
        abline(h = seq(0.1,1,.1), lty = 2)
        title(main=paste('\nbacklog:',backlog,'; HP_miner:',miner.hashpower),outer=TRUE)
        par(new=FALSE)}
      }
    win.chances.all<-c(win.chances.all,win.chances.last)
    table.results[as.character(100-miner.hashpower),as.character(backlog)]<-win.chances.last #save result in matrix
  }
  
  #OVERVIEW
  if (length(breaks.all!=0) & !table.only) { #
    plot.new
    par(mfrow=c(1,1))
    IX.H<-seq(3,by=6,length.out=length(BL)) # max. number of H's
    IX.H<-IX.H[IX.H<length(H.all)] #limited to actual number of H's
  
    breaks.max<-max(breaks.all)
    win.chances.max<-max(win.chances.all)
    X<-c(0,breaks.max)
    Y<-c(0,win.chances.max)
    plot(X,Y,xlab='time',ylab='win chance',xlim=X,ylim=Y,type='n') #setup frame
  
    legend.string<-legend.kleur<-c()
    ix.wc<-0
    for (ix.H in IX.H){
      ix.wc<-ix.wc+1
      points(H.all[[ix.H]][1:length(H.all[[ix.H]])],H.all[[ix.H+1]],col=H.all[[ix.H+2]],type='s',xlab='',ylab='')
      abline(h = seq(0,1,.1), lty = 2)
      legend.string<-c(legend.string,H.all[[ix.H-1]])
      legend.kleur<-c(legend.kleur,H.all[[ix.H+2]])
    }
    legend('topright',legend=legend.string,col=legend.kleur,lwd=2,title='backlog')
    title(main=paste('\n\nWIN CHANCE ATTACKER\nHP_miner:',H.all[[1]]),outer=TRUE)
  }
}
```

```{r,eval=FALSE}
for (miner.hashpower in HPM){

  H.all<-list()
  ix.kleur<-0
  win.chances.all<-breaks.all<-c()
  
  for (backlog in BL){

    par(mfrow=c(2,2))
    par(new=FALSE)
    ix.kleur<-ix.kleur+1
    min.time <- backlog # note that the attacker can not win by definition if time<backlog
    rep.factor <- max.time-min.time+1 # technical parameter to drive the lengths of the vectors for mcmapply
    res<-mcmapply(simulate,rep(chains,rep.factor),rep(backlog,rep.factor),seq(min.time,max.time),rep(miner.hashpower,rep.factor),rep(iterations,rep.factor),SIMPLIFY = 'array')
    wins<-res[1,]
    catch.up<-res[2,]
    
    if (all(is.na(catch.up))) {
      win.chances.last<-0
    } else {
      # er is tenminste een keer een catchup geweest
          
      #PLOT 1

      H<-hist(catch.up,breaks=50,main='HISTOGRAM CATCHUP TIMES',xlab='time',ylab='freq',col='red',plot=!table.only)
      win.chances<-cumsum(H$counts)/(rep.factor)
      win.chances.last<-win.chances[length(win.chances)]

      if (!table.only){

        H.all<-c(H.all,list(miner.hashpower,backlog,H$breaks[2:length(H$breaks)],win.chances,kleuren[ix.kleur],sum(wins))) #save info
    
        #PLOT 2
        plot(H$breaks[2:length(H$breaks)],win.chances,main='CUM CHANCE CATCHUP TIMES',xlab='time',ylab='cum chance',col='red',type='h')
        abline(h = seq(0.1,1,.1), lty = 2)

        breaks.last<-H$breaks[length(H$breaks)]
        breaks.all<-c(breaks.all,breaks.last) #for overview usage
    
        #PLOT 3
        plot(H$breaks[2:length(H$breaks)],win.chances,main='WIN CHANCES',xlab='time',ylab='win.chances',col='green',type='s')
        abline(h = seq(0.1,1,.1), lty = 2)
        title(main=paste('\nbacklog:',backlog,'; HP_miner:',miner.hashpower),outer=TRUE)
        par(new=FALSE)}
      }
    win.chances.all<-c(win.chances.all,win.chances.last)
    table.results[as.character(100-miner.hashpower),as.character(backlog)]<-win.chances.last #save result in matrix
  }
  
  #OVERVIEW
  if (length(breaks.all!=0) & !table.only) { #
    plot.new
    par(mfrow=c(1,1))
    IX.H<-seq(3,by=6,length.out=length(BL)) # max. number of H's
    IX.H<-IX.H[IX.H<length(H.all)] #limited to actual number of H's
  
    breaks.max<-max(breaks.all)
    win.chances.max<-max(win.chances.all)
    X<-c(0,breaks.max)
    Y<-c(0,win.chances.max)
    plot(X,Y,xlab='time',ylab='win chance',xlim=X,ylim=Y,type='n') #setup frame
  
    legend.string<-legend.kleur<-c()
    ix.wc<-0
    for (ix.H in IX.H){
      ix.wc<-ix.wc+1
      points(H.all[[ix.H]][1:length(H.all[[ix.H]])],H.all[[ix.H+1]],col=H.all[[ix.H+2]],type='s',xlab='',ylab='')
      abline(h = seq(0,1,.1), lty = 2)
      legend.string<-c(legend.string,H.all[[ix.H-1]])
      legend.kleur<-c(legend.kleur,H.all[[ix.H+2]])
    }
    legend('topright',legend=legend.string,col=legend.kleur,lwd=2,title='backlog')
    title(main=paste('\n\nWIN CHANCE ATTACKER\nHP_miner:',H.all[[1]]),outer=TRUE)
  }
}
```

`r paste('CONVERGENCE TABLE (',iterations,'ITERATIONS )')`
```{r show table,echo=FALSE}
print(table.results)
```

`r "rows contain haspower.attacker (Rosenfeld's q) and columns contain backlog values"`

`r 'simulation time:'`
```{r show simulation time,echo=FALSE}
time.end<-proc.time()
print(time.end-time.start)
```
